---
title: "R Notebook"
output: html_notebook
---

```{r}
setwd("../../data/clean")

topic_loc    <- read_feather('../../data/clean/location_topic_weights.feather')
topic_artist <- read_feather('../../data/clean/artist_topic_weights.feather')
topic_year   <- read_feather('../../data/clean/year_topic_weights.feather')
```



When there are less songs/year, the topics are less consistant.
```{r}
df_parallel_coord <- data.frame(t(topic_year))[1:3,]
colnames(df_parallel_coord) <- topic_year$year
df_parallel_coord$topic<-rownames(df_parallel_coord)

GGally::ggparcoord(df_parallel_coord, scale='globalminmax', columns = 1:87, groupColumn='topic')+ theme(text = element_text(size=5),axis.text.x = element_text(angle=90, hjust=1))
```

Considering years once there are more songs reeased, there are no clear trends.
```{r}
df_parallel_coord <- data.frame(t(topic_year))[1:3,]
colnames(df_parallel_coord) <- topic_year$year
df_parallel_coord$topic<-rownames(df_parallel_coord)

GGally::ggparcoord(df_parallel_coord, scale='globalminmax', columns = 57:87, groupColumn='topic')+ theme(text = element_text(size=15),axis.text.x = element_text(angle=90, hjust=1))
```

* Results show where a lot of music is made (usa, uk)
* Results show that relgion  defines an artist far more than death, or love - more artists have a religion score over 80 than any other topic
* Results dont indicate a particular region which has more of one genre


Clearly most music in the dataset is made in the United States and in Europe.
```{r}
topic_loc_tidy <- gather(topic_loc, 'topic', 'weight', 2:4)
topic_loc_tidy$topic <- as.factor(topic_loc_tidy$topic)

ggplot(topic_loc_tidy) +
  borders("world", colour="gray50", fill="gray90") +
  geom_point(aes(x=longitude, y=latitude) , color = 'red', size=0.1, alpha=.25)+
  ggtitle('Topics by Location')

```

Most "topic-heavy" artists: their average song has a score in the topic over 0.9
```{r}

data_to_plot <- select(topic_loc_tidy, longitude, latitude, topic, weight) %>% filter(weight>0.75)
ggplot(data_to_plot) +
  borders("usa", colour="gray50", fill="gray90") +
  geom_point(aes(x=longitude, y=latitude, color=topic, shape=topic) , size=2, alpha=.75)+
  xlim(-125, -65)+
  ylim(25,50)+
  ggtitle('Topics in USA')
```

"topic heavy" plot in Europe 
```{r}
ggplot(data_to_plot) +
  borders("world", colour="gray50", fill="gray90") +
  geom_point(aes(x=longitude, y=latitude, color=topic) , size=2, alpha=.3)+
  xlim(-25, 50)+
  ylim(25,75)+
  ggtitle('Europe')
```
  
"topic heavy" plot in UK 
```{r}
ggplot(data_to_plot) +
  borders("world", colour="gray50", fill="gray90") +
  geom_point(aes(x=longitude, y=latitude, color=topic) , size=2, alpha=.75)+
  xlim(-12.5, 5)+
  ylim(48,60)+
  ggtitle('United Kingdom')
```

similar numbers of each topic - thats not visible because many of them are the same artists, and the blue dot ends up on top of the same artis
```{r}
data_to_plot <- select(topic_loc_tidy, longitude, latitude, topic, weight) %>% filter(weight>0.90)
ggplot(data_to_plot, aes(topic))+geom_bar()
```



