# 3. Data Quality

```{r}
library(feather)
library(tidyverse)

df_mxm_dataset <-      read_feather("../data/transform/mxm_dataset.feather")
df_top_1000_dataset <- read_feather("../data/transform/bag_of_words_top_1000.feather")
df_artist_topic <-     read_feather("../data/transform/artist_topic_weights.feather")
df_year_topic <-       read_feather("../data/transform/year_topic_weights.feather")
```

## 3.1. Missing Data

## 3.2. Song Word Count

Per song counts must be carefully considered, because they only count words which are in the most common 5000 words. Nevertheless, we can see that the mode is around 75 words, and there is a wide range of word counts.

```{r}
mxm_dup <- df_mxm_dataset
mxm_dup$track_id <-NULL
mxm_dup <- transform(mxm_dup, sum=rowSums(mxm_dup))
word_count <- data.frame(df_mxm_dataset$track_id, mxm_dup$sum)

ggplot(word_count, aes(mxm_dup.sum)) +
  geom_histogram(binwidth=25)+
  scale_x_continuous(limits = c(0,1250))+
  ggtitle('Song Word Counts')+
  xlab('Word Count')+
  ylab('Song Count')

ggplot(word_count, aes(mxm_dup.sum)) +
  geom_histogram(binwidth=1)+
  scale_x_continuous(limits = c(0,100))+
  ggtitle('Song Word Counts (low word count)')+
  xlab('Word Count')+
  ylab('Song Count')
```


## 3.3. Song Count Over Time

The number of songs/year grows exponentially in size until ~2005, before it slows in growth rate and then declines.

```{r}
# songs/year distrubution
ggplot(df_year_topic,aes(x=year, y=song_count))+
  geom_point()+
  geom_line()+
  xlab('Year')+
  ylab('Song Count')+
  ggtitle('Song Count/Year')
```

## 3.4. Song Count per Artist

The dataset is skewed towards including few songs for each artist

```{r}
#songs/artist distribution
ggplot(df_artist_topic,aes(song_count))+
  geom_histogram(binwidth=1)+
  xlab('Song Count')+
  ylab('Number of Artists')+
  ggtitle('Song Count/Artist')
```



