---
title: "words_per_year"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reading words per year
```{r}
words_year <- readRDS(file = "words_per_year.Rda")
```

Plotting usage of 'love' and 'hate'
```{r}
library(ggplot2)

ggplot(words_year,aes(x=track_year)) + geom_line(aes(y = love, colour = 'love')) + geom_line(aes(y = hate, colour = 'hate')) +  ylim(0,20000) + xlim(1960,2010)

```

Plotting usage of 'jesus' and 'god'
```{r}
ggplot(words_year,aes(x=track_year)) + geom_line(aes(y = jesus, colour = 'jesus')) + geom_line(aes(y = god, colour = 'god')) +  ylim(0,3000) + xlim(1960,2010)

```

Plotting usage of 'guitar' and 'hate'
```{r}
ggplot(words_year,aes(x=track_year)) + geom_line(aes(y = guitar, colour = 'guitar'))  + geom_line(aes(y = hate, colour = 'hate')) +  ylim(0,1500) 

```


Reading tracks per year
```{r}
library(feather)
tracks_per_year <- read_feather('/Users/valmik/School/Spring\ 2018/EDAV/song-lyrics/data/transform/tracks_per_year.feather')
```

Reading mxm data
```{r}
mxm <- read_feather('/Users/valmik/School/Spring\ 2018/EDAV/song-lyrics/data/transform/mxm_dataset.feather')
```

Filtering tracks to include only those present in mxm
```{r}
tracks_per_year <- tracks_per_year[tracks_per_year$track_id %in% mxm$track_id,]
```

Gettings number of songs for each year
```{r}
library(dplyr)
songs_year <- data.frame(table(tracks_per_year$year))
colnames(songs_year)[1] <- 'track_year'
```

Adding number of songs to words per year
```{r}
words_year <- merge(words_year, songs_year)
```

Convertings to appearance of words per song
```{r}
cols = colnames(words_year)
remove = c('track_year', 'Freq')
cols <- cols[! cols %in% remove]
```

```{r}
words_year[cols] <- lapply(words_year[cols], '/', words_year$Freq)
```

Plotting the same plots again with standardized values
```{r}
library(ggplot2)

ggplot(words_year,aes(x=track_year)) + geom_line(aes(y = love, colour = 'love')) + geom_line(aes(y = hate, colour = 'hate')) +  ylim(0,4) + xlim(1960,2010)

```
```{r}
ggplot(words_year,aes(x=track_year)) + geom_line(aes(y = jesus, colour = 'jesus')) + geom_line(aes(y = god, colour = 'god')) +  ylim(0,0.3) + xlim(1960,2010)

```

```{r}
ggplot(words_year,aes(x=track_year)) + geom_line(aes(y = guitar, colour = 'guitar'))  + geom_line(aes(y = hate, colour = 'hate')) +  ylim(0,4) 

```

