---
title: "R Notebook"
output: html_notebook
---

# Distance-based comparisons

One of the things we are interested on is to find which observations are more similar/different to each other. Using the word embeddings representation of
our data, we computed pairwise euclidean distance between observations at different levels to find similar elements.

```{r}
library(feather)
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
source('experiments/eduardo/util.R')
```

```{r}
# load embeddings data data and
df <- read_feather('data/transform/embeddings.feather')

# get columns corresponding to the metadata words
columns <- colnames(df)
columns_metadata <- columns[endsWith(columns, '_')]
columns_lyrics <- columns[!endsWith(columns, '_')]
```

# Finding similar years


```{r}
by_year_en <- df %>% filter(language_ == 'en') %>%
                 group_by(release_year_) %>%
                 do(mean_words(., columns_lyrics))
plot_matrix(by_year_en)
```

```{r}
by_year_es <- df %>% filter(language_ == 'es') %>% group_by(release_year_)
                %>% do(mean_words(., columns_lyrics))
plot_matrix(by_year_es)
```


## Finding similar genres


```{r}
genre_ <- table(df$genre_)
top_genres <- names(genre_[genre_ > 80])

by_genre <- df %>% filter(genre_ %in% top_genres) %>% group_by(genre_) %>%
                do(mean_words(., columns_lyrics))

plot_matrix(by_genre)
```

```{r}
by_genre_es <- df %>% filter(genre_ %in% top_genres, language_ == 'es') %>%
                    group_by(genre_) %>% do(mean_words(., columns_lyrics))

plot_matrix(by_genre_es)
```

```{r}
by_genre_en <- df %>% filter(genre_ %in% top_genres, language_ == 'en') %>%
                    group_by(genre_) %>% do(mean_words(., columns_lyrics))

plot_matrix(by_genre_en)
```

```{r}
reggaeton <- df %>% filter(genre_ == 'Reggaeton')
table(reggaeton$language_)

reggaeton_en <- df %>% filter(genre_ == 'Reggaeton', language_ == 'en')

reggaeton_en[c('title_', 'artist_name_')]
```


## Finding similar artists

```{r}
reggaeton <- df %>% filter(genre_ == 'Reggaeton')

by_artist <- reggaeton %>% group_by(artist_name_) %>%
                do(mean_words(., columns_lyrics))

plot_matrix(by_artist)
```

```{r}
sub <- df %>% filter(genre_ == 'Rock') %>% sample_n(50)

unique(sub$artist_name_)

by_artist <- sub %>% group_by(artist_name_) %>% do(mean_words(., columns_lyrics))

plot_matrix(by_artist)
```


# Comparing songs form the same artist

```{r}
oasis <- df %>% filter(artist_name_ == 'Oasis') %>% sample_n(30)
oasis <- oasis[c('title_', columns_lyrics)]
plot_matrix(oasis)
```

```{r}
deep purple <- df %>% filter(artist_name_ == 'Deep Purple') %>% sample_n(30)
deep purple <- deep purple[c('title_', columns_lyrics)]
plot_matrix(deep purple)
```


## Find similar songs (bag of words and embeddings)

TODO: visualize artists in two dimensions using PCA

## Songs that talk about the same topic